### 2014-11-09 implementation.tex
Integrating Gustavo\'s edits

\name is implemented as a shallow extension of GHC Haskell and runs on top of
Cassandra, an off-the-shelf eventually consistent distributed data (or backing)
store responsible for all data management issues (i.e., replication, fault
tolerance, availability, and convergence).  Template Haskell is used implement
static contract classification, and proof obligations are discharged with the
help of the Z3~\cite{Z3} SMT solver. Figure~\ref{fig:impl_mod} illustrates the
overall system architecture.

>>>>

\name is implemented as a shallow extension of GHC Haskell and runs on top of
Cassandra, an off-the-shelf eventually consistent distributed data (or backing)
store responsible for all data management issues (i.e., replication, fault
tolerance, availability, and convergence).  Template Haskell is used to
implement static contract classification, and proof obligations are discharged
with the help of the Z3~\cite{Z3} SMT solver. Figure~\ref{fig:impl_mod}
illustrates the overall system architecture.

### 2014-11-09 intro.tex
Integrating Gustavo\'s edits

Many real-world web services --- such as those built and maintained by Amazon,
Facebook, Google, Twitter, etc. --- replicate application state and logic
across multiple \emph{replicas} within and across data centers. Replication is
intended not only to improve application throughput and reduce user-perceived
latency, but also to tolerate partial failures without compromising overall
service availability. Traditionally programmers have relied on \emph{strong
consistency} guarantees such as linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct applications.
While strong consistency is an easily stated property, it masks the reality
underlying large-scale distributed systems with respect to non-uniform latency,
availability and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
modern web services, which aim to provide an "always on" experience,
overwhelmingly favour availability and partition tolerance over strong
consistency. To this end, several \emph{weak consistency} models such as
eventual consistency, causal consistency, session guarantees, and timeline
consistency have been proposed.

>>>>

Many real-world web services --- such as those built and maintained by Amazon,
Facebook, Google, Twitter, etc. --- replicate application state and logic
across multiple \emph{replicas} within and across data centers. Replication is
intended not only to improve application throughput and reduce user-perceived
latency, but also to tolerate partial failures without compromising overall
service availability. Traditionally programmers have relied on \emph{strong
consistency} guarantees such as linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct applications.
While strong consistency is an easily stated property, it masks the reality
underlying large-scale distributed systems with respect to non-uniform latency,
availability and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
modern web services, which aim to provide an "always-on" experience,
overwhelmingly favour availability and partition tolerance over strong
consistency. To this end, several \emph{weak consistency} models such as
eventual consistency, causal consistency, session guarantees, and timeline
consistency have been proposed.

### 2014-11-09 motivation.tex
edits
Integrating Gustavo\'s edits

\begin{figure}
\centering
\subfigure[Unsafe withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
\hfill
\subfigure[Negative balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
\hfill
\subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
\caption{Anomalies possible under eventual consistency for the get balance operation.}
\label{fig:cleanliness_examples}
\end{figure}

@@

It is easy to see that preventing concurrent \cf{withdraw} operations
eliminates this anamoly. This can be done by insisting that \cf{withdraw} be
executed as a strongly consistent operation. Despite this strengthening,
\cf{getBalance} operation may incorrectly report a negative balance to the end
user. Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly},
which consists of three concurrent sessions performing a \cf{deposit}, a
\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same bank
account object. As the \cf{vis} edge indicates, operation \cf{withdraw(50)} in
session 2, witnesses the effects of \cf{deposit(100)} from session 1, concludes
that there is sufficient balance, and completes successfully. However, the
\cf{getBalance} operation may only witness this successful withdraw, but not
the causally preceding \cf{dopsit}, and reports the balance of negative 50 to
the user.

@@

\name helps facilitate the mapping of operations to appropriate consistency
levels by letting the programmer declare application-level consistency
constraints as \emph{contracts} (Fig~\ref{fig:contract-lang}\footnote{\name
exposes the contract construction language as a Haskell library}) that
axiomatically specify the set of allowed executions involving this operation.
In the case of the bank account, any execution that does not exhibit the
anomalies described in the previous section is a \emph{well-formed} execution
on the bank account object. By specifying the set of legal executions for each
data type in terms of a trace of operation invocations on that type, \name\
ensures that all executions over that type are well-formed.

@@

\vspace{-1em}
\begin{smathpar}
\begin{array}{l}
\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
\end{array}
\end{smathpar}

>>>>

\begin{figure}
\centering
\subfigure[Unsafe withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
\hfill
\subfigure[Negative balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
\hfill
\subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
\caption{Anomalies possible under eventual consistency for the get balance operation.}
\label{fig:ba_anomalies}
\end{figure}

@@

It is easy to see that preventing concurrent \cf{withdraw} operations
eliminates this anamoly. This can be done by insisting that \cf{withdraw} be
executed as a strongly consistent operation. Despite this strengthening,
\cf{getBalance} operation may still incorrectly report a negative balance to
the end user. Consider the execution shown in
fig.~\ref{fig:negativeBalanceAnomaly}, which consists of three concurrent
sessions performing a \cf{deposit}, a \cf{withdraw}, and a \cf{getBalance}
operation, respectively, on the same bank account object. As the \cf{vis} edge
indicates, operation \cf{withdraw(50)} in session 2, witnesses the effects of
\cf{deposit(100)} from session 1, concludes that there is sufficient balance,
and completes successfully. However, the \cf{getBalance} operation may only
witness this successful withdraw, but not the causally preceding \cf{dopsit},
and reports the balance of negative 50 to the user.

@@

\name helps facilitate the mapping of operations to appropriate consistency
levels by letting the programmer declare application-level consistency
constraints as \emph{contracts}\footnote{\name exposes the contract
construction language as a Haskell library} (Figure~\ref{fig:contract-lang})
that axiomatically specify the set of allowed executions involving this
operation. In the case of the bank account, any execution that does not exhibit
the anomalies described in the previous section is a \emph{well-formed}
execution on the bank account object. By specifying the set of legal executions
for each data type in terms of a trace of operation invocations on that type,
\name\ ensures that all executions over that type are well-formed.

@@

\vspace{-1em}
\begin{smathpar}
\begin{array}{l}
\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
\qquad (\vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}) \\
\qquad \wedge~ ((\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff})
\end{array}
\end{smathpar}
